### 2025-01-06 reviewer-response.md

*On line 141, there is a function defined for the local fractal dimension
(LFD), but it is not expressed in functional notation.  In other words, it
should start with $LFD(X,r_1,r_2)$, since it is clearly a function of all
three.  This functional notation should propagate throughout the rest of
the manuscript.  Much the same is true on line 142, where it should read
$B(X,q,r)$, not $B_X(q,r)$.  If makes little sense to put $X$ as a
subscript in this context.*

similar to complaint from reviewer 2, address with same language

*In the algorithms, starting with Algorithm 2.1, the pseudocode style is
ugly.  Much better would be the pseudocode style from the fourth edition
of Introduction to Algorithms, CLRS.  Also, the description of $criteria$
is as a "stopping criteria", while the logic of the code has it as a
continuing criteria.  Please fix the pseudocode throughout.*

We evaluated the CLRS package, but did not see any aesthetic differences. We have, however, increased the font size for the algorithm environments, so that it should be more legible.
call it continuation criteria (NI)

### 2025-01-06 sections/4-results.tex

Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
Glove-25 under cosine distance, and
Sift under Euclidean distance.
Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with the same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.
The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
The left-most point on each line is at the cardinality of the original dataset.
The vertical axis denotes throughput in queries per second.
Both axes are on a logarithmic scale.
In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k = 100$ can be found in the Supplement.
For HNSW and ANNOY, we report the recall for each measurement in the plots. For algorithms which exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.

@@

Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset. For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
Which one of the three CAKES algorithms is fastest also differs by dataset.
For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset.
With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).

In contrast with the results on the ANN Benchmark datasets reported above, with the Random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly.
As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.

>>>>

Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
Glove-25 under cosine distance, and
Sift under Euclidean distance.
Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with the same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.
The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
The left-most point on each line is at the cardinality of the original dataset.
The vertical axis denotes throughput in queries per second.
Both axes are on a logarithmic scale.
In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k = 100$ can be found in the Supplement.
For HNSW and ANNOY, we report the recall for each measurement in the plots. For algorithms that exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.

@@

Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset. For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively.
Which one of the three CAKES algorithms is fastest also differs by dataset.
For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve.
On all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance that is \textit{constant} in the cardinality of the dataset.
With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling.
We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink).
For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$.
On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY degrades with cardinality.
At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift.
In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (where the distance function is a metric), and near-perfect recall on Glove-25 (cosine distance is not a metric).

In contrast with the results on the ANN Benchmark datasets reported above, as seen in Figure~\ref{fig:results:random-scaling}, we observe performs quite slowly pn the Random dataset.
As with the real datasets, HNSW and ANNOY are the fastest algorithms, and CAKES exhibits perfect recall at all cardinalities.
HNSW and ANNOY exhibit \textit{much} lower recall on this random dataset than on any of the ANN benchmark datasets;
in particular, with a multiplier of 1, HNSW and ANNOY have recall as low as 0.060 and 0.028 respectively, as reported in Tables~\ref{tab:results:qps-and-recall-fmn},~\ref{tab:results:qps-and-recall-glove},~\ref{tab:results:qps-and-recall-sift}, and~\ref{tab:results:qps-and-recall-random}.

### 2025-01-06 sections/5-discussion.tex

When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{\lceil d \rceil} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD,
the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
This suggests that rather than using the reciprocal of the \textit{harmonic} mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.

>>>>

When written with the same notation as used in Section~\ref{sec:methods:knn-search:complexity-of-sieve-methods}, we see that the time complexity of Repeated $\rho$-NN is $\mathcal{O}(\tfrac{\mathcal{T}}{\lceil d \rceil} + \mathcal{L})$, where $\mathcal{T}$ is the time complexity of tree-search and $\mathcal{L}$ is the time complexity of leaf-search.
Even though Repeated $\rho$-NN has the lowest time complexity (compare to $\mathcal{O}(\mathcal{T}\log{(\mathcal{T}+\mathcal{L})} + \mathcal{L}\log{(\mathcal{T}+\mathcal{L})})$ for Breadth-First Sieve and $\mathcal{O}(\mathcal{T}\log{\mathcal{T}} + \mathcal{L}\log{k})$ for Depth-First Sieve) of the three CAKES algorithms, it is not always the fastest algorithm empirically.
We believe that some of this discrepancy can be explained by the fact that if Repeated $\rho$-NN significantly ``overshoots'' the correct radius for $k$ hits ($\rho_k$) during tree-search, leaf-search will require looking at more points, rendering the true scaling factor higher than that in Equation~\ref{eq:methods:repeated-rnn-complexity}.
This overshot can occur when the LFDs of clusters near the query are not concentrated around their expectation.
For example, if most clusters near the query have very low LFD except for one anomaly with very high LFD,
the harmonic mean LFD $\mu$ can still be low, so the factor of radial increase in (Equation)~\ref{eq:methods:repeated-rnn-factor} may be much larger than necessary for guaranteeing $k$ hits.
This suggests that rather than using the reciprocal of the harmonic mean LFD in~\ref{eq:methods:repeated-rnn-factor}, we may achieve better results with a mean that is more sensitive to high outliers, such as the geometric mean.
We leave it as an avenue for future work to characterize when Repeated $\rho$-NN significantly ``overshoots'' $\rho_k$ and to improve upon the factor of radial increase in~\ref{eq:methods:repeated-rnn-factor} so that this occurs less frequently and with less severity.
